Article Title

### 2013-10-01 sections/Regularization.tex
clarified "Evaluation of estimators" based on Kreso's input.

\subsection*{Data acquisition}
We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.  Numerious repetitions of full-field drifting gratings (Fig. 1A and 1B) were presented to the eye contralateral to the imaged site. This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4. Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).  The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).

\input{sections/Figure1.tex}

\subsection*{The sample covariance matrix}
We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons. Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.

For more rigorous notation, definitions, and derivations, see Appendix.

The usual estimator of $\Sigma$ is the sample covariance matrix
\begin{equation}
\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
\end{equation}
where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom. For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom. When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.

The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.

\subsection*{Evaluation of covariance matrix estimates}
The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.

For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
\begin{equation}\label{eq:loss}
\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
\end{equation}
This choice is motivated by mathematical convenience. Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.  We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.

The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
\begin{equation}
r = \E{\loss{\hat\Sigma, \Sigma}}
\end{equation}

In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
\begin{equation}
\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
\end{equation}
 and its expected value or \emph{empirical risk} is
\begin{equation}\label{eq:empiricalRisk}
\hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
\end{equation}
Because the chosen loss function is linear in its second argument in the sense that
\begin{equation}
\loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
\end{equation}
the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
\begin{equation}
\hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
\end{equation}
Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.

Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.

\subsection*{Regularization}
Under many loss functions\footnote{
The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$. This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. The bias-variance decomposition is presented here to motivate the use of regularization.}, 
the estimator's risk can be decomposed as the sum
\begin{equation}\label{eq:bias-variance}
r = b + \varepsilon
\end{equation}
of \emph{approximation error} (``bias'' or systematic error)
\begin{equation}
b = \loss{\bar\Sigma,\Sigma}
\end{equation}
and \emph{estimation error} (``variance'') 
\begin{equation}
\varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
\end{equation}
where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.

@@

\input{sections/Estimators.tex}

\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
\input{sections/Figure3.tex}

\subsection*{In visual cortex, Estimator D dominates}
\input{sections/Figure4.tex}
This is the full matrix of all comparisons of the estimators.  The plots represents the histograms of the differences of the loss function between the pair of estimators for all 31 sites.  When the differences are positive, the first estimate in the title outperforms the second estimate.

Of particular interest is the first row, which shows that all regularized estimates  outperformed the sample covariance estimate.  Even more informative is the last column, which shows that estimate D (sparse+lowrank) significantly outperformed all other estimates.

>>>>

\subsection*{Regularization}
Under many loss functions\footnote{
The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. 
However, the equality does hold for its close cousin, Stein's \emph{entropy loss}, which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$. 
This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. 
The bias-variance decomposition is presented here to motivate the use of regularization.}, 
the estimator's risk can be decomposed as the sum
\begin{equation}\label{eq:bias-variance}
    r = b + \varepsilon
\end{equation}
of \emph{approximation error} (``bias'' or systematic error)
\begin{equation}
   b = \loss{\bar\Sigma,\Sigma}
\end{equation}
and \emph{estimation error} (``variance'') 
\begin{equation}
   \varepsilon = \E[\hat\Sigma]{\loss{\hat\Sigma, \bar\Sigma}}
\end{equation}
where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.

@@

### 2013-10-01 sections/Theory.tex
clarified "Evaluation of estimators" based on Kreso's input.

\subsection*{Evaluation of covariance matrix estimates}
The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.

For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
\begin{equation}\label{eq:loss}
\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
\end{equation}
This choice is motivated by mathematical convenience. Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.  We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.

The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
\begin{equation}
r = \E{\loss{\hat\Sigma, \Sigma}}
\end{equation}

In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
\begin{equation}
\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
\end{equation}
 and its expected value or \emph{empirical risk} is
 \begin{equation}\label{eq:empiricalRisk}
 \hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
 \end{equation}
 Because the chosen loss function is linear in its second argument in the sense that
 \begin{equation}
 \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
 \end{equation}
 the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
 \begin{equation}
 \hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
 \end{equation}
 Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.

 Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.

 \subsection*{Regularization}
 Under many loss functions\footnote{
 The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$. This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. The bias-variance decomposition is presented here to motivate the use of regularization.}, 
 the estimator's risk can be decomposed as the sum
 \begin{equation}\label{eq:bias-variance}
 r = b + \varepsilon
 \end{equation}
 of \emph{approximation error} (``bias'' or systematic error)
 \begin{equation}
 b = \loss{\bar\Sigma,\Sigma}
 \end{equation}
 and \emph{estimation error} (``variance'') 
 \begin{equation}
 \varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
 \end{equation}
 where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.

 The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.

 The estimator risk can be reduced by \emph{regularization}. Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}. A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.  Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.  Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}. Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.

>>>>

\subsection*{Evaluation of covariance matrix estimators}
The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.

For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
\begin{equation}\label{eq:loss}
\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
\end{equation}
This choice is motivated by mathematical convenience. Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.  We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.

The aim of our project is to produce covariance matrix estimates that minimize the expected loss 
\begin{equation}
r = \E{\loss{\hat\Sigma, \Sigma}}
\end{equation}
which is know as the \emph{risk} of $\hat\Sigma$.

In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through \emph{validation}. 
Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.

$\loss{\cdot,\cdot}$ is additive in its second argument such that
 \begin{equation}\label{eq:additivity}
 \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
 \end{equation}
Then \emph{validation loss}  \TODO{check appropriate term}
\begin{equation}
\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
\end{equation}
is an unbiased estimate of the risk:
 \begin{equation}\label{eq:empiricalRisk}
\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
= \E{\loss{\hat\Sigma,\Sigma}} = r
 \end{equation}
Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.

Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.

Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation leads to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
